Computing Entropies With Nested Sampling

### 2017-07-11 paper/ms.tex
Pareto stuffs

\renewcommand{\d}{\boldsymbol{d}}
\newcommand{\todo}{\color{orange} \bf}
\newcommand{\query}{\color{green} \bf}
\newcommand{\x}{\boldsymbol{\theta}}
\newcommand{\bphi}{\boldsymbol{\phi}}
\newcommand{\boldeta}{\boldsymbol{\eta}}
\newcommand{\n}{\boldsymbol{\eta}}
\newcommand{\depth}{(\textnormal{depth})}
\newcommand{\xref}{\x_{\rm ref}}

@@

Consider $n=100$ data values, partitioned into two halves, so
$\boldsymbol{y} = \{x_1, ..., x_{50}\} = \{y_1, ..., y_{50}\}$
is the first half of the data and
$\boldsymbol{z} = \{x_{51}, ..., x_{100}\} = \{z_1, ..., z_{50}\}$
is the second half. Let $\bar{y}$ and $\bar{z}$ be the averages of the two
subsets of the data:
\begin{align}
\bar{y} &= \frac{1}{50} \sum_{i=1}^{50} y_i \\
\bar{z} &= \frac{1}{50} \sum_{i=1}^{50} z_i.
\end{align}
In this section we use the algorithm to answer the questions:
i) How uncertain is $\bar{y}$?; ii) How uncertain is $\bar{z}$; and
iii) How informative is $\bar{y}$ about $\bar{z}$? These are quantified by
$H(\bar{y})$, $H(\bar{z})$, and $I(\bar{y}; \bar{z})$ respectively. This can
be done despite the fact that $\alpha$ is unknown.

The prior for $\alpha$ was chosen to be lognormal such that the prior median
of $\alpha$ is 1 and the prior standard deviation of $\ln \alpha$ is also 1.
Due to the heavy-tailed nature of the Pareto distribution, I ran the
algorithm in terms of $\ln \bar{y}$ and $\ln \bar{z}$. This affects the
differential entropies but not the mutual information (which is invariant 
under changes of parameterisation).

To compute the entropy of $\ln \bar{y}$, the distance function was just the
absolute value of the distance between the NS particles and the
reference particle in terms of $\ln \bar{y}$.

>>>>

\renewcommand{\d}{\boldsymbol{d}}
\newcommand{\todo}{\color{orange} \bf}
\newcommand{\query}{\color{green} \bf}
\newcommand{\x}{\boldsymbol{\theta}}
\newcommand{\bphi}{\boldsymbol{\phi}}
\newcommand{\boldeta}{\boldsymbol{\eta}}
\newcommand{\n}{\boldsymbol{\eta}}
\newcommand{\depth}{(\textnormal{depth})}
\newcommand{\xref}{\x_{\rm ref}}
\newcommand{\ytot}{y_{\rm tot}}
\newcommand{\ztot}{z_{\rm tot}}

@@

Consider $n=100$ data values, partitioned into two halves, so
$\boldsymbol{y} = \{x_1, ..., x_{50}\} = \{y_1, ..., y_{50}\}$
is the first half of the data and
$\boldsymbol{z} = \{x_{51}, ..., x_{100}\} = \{z_1, ..., z_{50}\}$
is the second half. Let $\ytot$ and $\ztot$ be the averages of the two
subsets of the data:
\begin{align}
\ytot &= \sum_{i=1}^{50} y_i \\
\ztot &= \sum_{i=1}^{50} z_i.
\end{align}
In this section we use the algorithm to answer the questions:
i) How uncertain is $\ytot$?; ii) How uncertain is $\ztot$; and
iii) How informative is $\ytot$ about $\ztot$? These are quantified by
$H(\ytot)$, $H(\ztot)$, and $I(\ytot; \ztot)$ respectively. This can
be done despite the fact that $\alpha$ is unknown.

The prior for $\alpha$ was chosen to be lognormal such that the prior median
of $\alpha$ is 1 and the prior standard deviation of $\ln \alpha$ is also 1.
Due to the heavy-tailed nature of the Pareto distribution, I ran the
algorithm in terms of $\ln \ytot$ and $\ln \ztot$. This affects the
differential entropies but not the mutual information (which is invariant 
under changes of parameterisation).

\begin{figure}[!ht]
\centering
\includegraphics[width=0.6\textwidth]{figures/pareto.pdf}
\caption{The joint distribution for the log-sum of the first half of the
`Pareto data' and the second half. The goal is to calculate the entropy of
the marginal distributions, the entropy of the joint distribution, and
hence the mutual information.\label{fig:pareto}}
\end{figure}

To compute the entropy of $\ln \ytot$, the distance function was just the
absolute value of the distance between the NS particles and the
reference particle in terms of $\ln \ytot$.
